---
title: "Assignment 4"
author: "Ana Zhibaj"
output: html_document
---

# Black Lives Matter protests in Boston

I want to visualize the spatial impact the Black Lives Matter protests had in Boston, particularly thinking about people following them from the windows of their apartment buildings, unable to physically be there. The following text from Johanna Hedva's Sick Woman Theory (2016) inspired me:

"I started to think about what modes of protest are aforded to sick people - it seemed to me that many people for whom Black Lives Matter is especially in service, might not be able to be present for the marches because they were imprisoned by a job, the threat of being fired from their job if they marched, or literal incarceration, and of course the threat of violence and police brutality - but also because of illness or disability, or because they were caring for someone with an illness or disability.

I thought of all the other invisible bodies, with their fists up, tucked away and out of sight." 

I make the case that the sound of the protest is the device that engages these invisible bodies from their windows. I chose 5 minutes (driving and walking) as arbitrary measures of impact of the sound of the protests (BLM protests took many shapes, some were vigils that were quiet, others were loud, so using walking and driving as measures of minimum and maximum impact). 

```{r, message=FALSE}
library(osmdata)
library(opentripplanner)
library(tidyverse)
library(sf)
library(ggthemes)
library(ggspatial)
library(osmar)
library(raster)
library(tidygeocoder)
library(ggplot2)
library(ggthemes)
library(ggalt)
```

The map layers I used are Black Lives Matters protests map from Alex Smith at Creosote maps and the buildings from analyze Boston. I converted the buildings into points, using the centroid function. 

```{r load map points, message=FALSE}
blm_protests <- st_read(
  "Alex_Smith_protest_points_2020_0929.shp")
```

```{r add buildings kml}
buildings <- st_read("http://bostonopendata-boston.opendata.arcgis.com/datasets/8bf3e3b0bde0432c82f76ee6a0608e7d_0.kml?outSR=%7B%22latestWkid%22%3A2249%2C%22wkid%22%3A102686%7D", quiet = TRUE)
```


```{r project on MA state plane}
MA_state_plane <- "+proj=lcc +lat_1=41.71666666666667 +lat_2=42.68333333333333 +lat_0=41 +lon_0=-71.5 +x_0=200000 +y_0=750000 +ellps=GRS80 +units=m +no_defs"

boston_street_features <- opq(bbox = 'Boston MA USA') %>%
  add_osm_feature(key = 'highway') %>%
  osmdata_sf()

boston_streets <- boston_street_features$osm_lines %>%
  st_transform(crs = MA_state_plane)
```

```{r transform buildings}
buildings <- buildings %>%
  st_transform(MA_state_plane)
```


```{r building centroids}
buildings <- st_centroid(buildings)
```

```{r, message=FALSE, results='hide'}
path_data <- file.path(getwd(), "OTP")
path_otp <- paste(path_data, "otp.jar",sep = "/")

otp_build_graph(otp = path_otp, dir = path_data, memory = 1024) 
```

```{r}
otp_setup(otp = path_otp, dir = path_data, memory =1024)
```

```{r}
# Connect to opentripplanner
otpcon <- otp_connect()
```

```{r, message=FALSE, results='hide'}
iso_5min_walk <- 
  otp_isochrone(otpcon = otpcon, fromPlace = blm_protests, 
                mode = "WALK", cutoffSec = 300) %>%
  st_transform(crs = MA_state_plane) %>%
  mutate(mode = "walk")

iso_5min_drive <- 
  otp_isochrone(otpcon = otpcon, fromPlace = blm_protests, 
                mode = "CAR", cutoffSec = 300) %>%
  st_transform(crs = MA_state_plane) %>%
  mutate(mode = "drive")

iso_all_modes <- rbind(iso_5min_drive, iso_5min_walk)

otp_stop()
```


```{r}
right_side <- st_bbox(iso_all_modes)$xmax
left_side  <- st_bbox(iso_all_modes)$xmin
top_side <- st_bbox(iso_all_modes)$ymax
bottom_side <- st_bbox(iso_all_modes)$ymin

ggplot(iso_all_modes) +
  annotation_map_tile(zoomin = 0, type = "stamenbw", progress = "none") +
  geom_sf(aes(fill = mode), alpha = 0.5) +
  geom_sf(data = blm_protests) +
  coord_sf(xlim = c(left_side, right_side), 
           ylim = c(bottom_side, top_side), expand = FALSE) +
  scale_fill_viridis_d(name = "Area that is reachable within 5 minutes",
                       labels = c("By car", "By foot")) +
  theme_map() +
  labs(caption = "Basemap Copyright OpenStreetMap contributors")
```



```{r}
iso_areas <- iso_all_modes %>%
  mutate(area = st_area(iso_all_modes)) %>%
  st_set_geometry(NULL) %>%
  pivot_wider(names_from = mode, values_from = area) 

ggplot(iso_areas, 
       aes(x = as.numeric(walk), y = as.numeric(drive))) +
  geom_point() +
  scale_x_continuous(name = 
            "Area within a five-minute walking distance\nof the protests\n(square km)",
            breaks = breaks <- seq(10000, 130000, by = 20000),
            labels = breaks / 1000000) +
  scale_y_continuous(name = 
            "Area within a five-minute driving distance\nof the protests\n(square km)",
            breaks = breaks <- seq(0, 1500000, by = 100000),
            labels = breaks / 1000000) +
  theme_bw()
```

As a last analysis, I wanted to calculate the number of buildings that are in the area of isochrones for all modes, but the map seems weird for some reason, and the numbers are way off. 

```{r subsetting trees in hubway stations}
buildings_blm <- buildings[iso_all_modes,]
ggplot(buildings_blm) +
  geom_sf() +
  geom_sf(data = buildings_blm, 
          color = "black", 
          size = 0.01) +
  theme_map()
```


```{r}
buildings <- buildings %>%
  st_join(buildings_blm) %>%
  mutate(by_blm = !is.na(Name.y))
```

How many buildings are there close to the sound of BLM protests?

```{r}
n_buildings_blm <- sum(buildings$by_blm)
n_buildings_blm
```

What percent of all buildings is that?

```{r}
n_buildings <- length(buildings$by_blm)
pct_buildings_blm <-n_buildings_blm / n_buildings
pct_buildings_blm
```










