---
title: "Cristina"
author: "Ana Zhibaj"
date: "10/3/2020"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r loading libraries, message=FALSE}
library(sf)
library(tidyverse)
library(ggthemes)
library(ggspatial)
library(units)
library(nngeo)
library(wesanderson)
library(extrafont)
library(extrafontdb)
library(units)
```

```{r, message=FALSE}
nhoods <- st_read("http://bostonopendata-boston.opendata.arcgis.com/datasets/3525b0ee6e6b427f9aab5d0a1d0a1a28_0.kml?outSR=%7B%22latestWkid%22%3A2249%2C%22wkid%22%3A102686%7D", quiet=TRUE)

trees <- st_read("http://bostonopendata-boston.opendata.arcgis.com/datasets/ce863d38db284efe83555caf8a832e2a_1.kml", quiet=TRUE)

openspaces <- st_read("http://bostonopendata-boston.opendata.arcgis.com/datasets/2868d370c55d4d458d4ae2224ef8cddd_7.kml", quiet=TRUE)

communityc <- st_read("http://bostonopendata-boston.opendata.arcgis.com/datasets/acedd06b39854088b0f2fefccffebf84_0.kml?outSR=%7B%22latestWkid%22%3A3857%2C%22wkid%22%3A102100%7D", quiet=TRUE)

water <- st_read("http://bostonopendata-boston.opendata.arcgis.com/datasets/2b3c0fa13b1c468eb702a3645fcc6bcb_5.kml", quiet=TRUE)

```

Since I'm going to incorporate areas and distances at least in one map, I'll transform my data to the **NAD83 / Massachusetts Mainland** coordinate system.

```{r, message=FALSE}
MA_NAD83_plane <- "+proj=lcc +lat_1=42.68333333333333 +lat_2=41.71666666666667 +lat_0=41 +lon_0=-71.5 +x_0=200000.0001016002 +y_0=750000 +ellps=GRS80 +datum=NAD83 +to_meter=0.3048006096012192 +no_defs"

nhoods <- nhoods %>%
  st_transform(MA_NAD83_plane)

trees <- trees %>%
  st_transform(MA_NAD83_plane)

openspaces <- openspaces %>%
  st_transform(MA_NAD83_plane)

communityc <- communityc %>%
  st_transform(MA_NAD83_plane)

water <- water %>%
  st_transform(MA_NAD83_plane)

```

------
BUFFER 

### Hubways: creating a Buffer 
400-meter buffer around each hubway stops in the City of Boston


```{r}
trees_buffer <- st_buffer(trees, dist = 250) %>%
  st_union()
ggplot(trees_buffer) + 
  geom_sf(color = "orange1") +
  theme_map()
```

```{r, message= FALSE}
trees_cc <- communityc[trees_buffer,]
ggplot(trees_cc) +
  geom_sf() +
  geom_sf(data = trees_cc, 
          color = "#bae4bc", 
          size = 0.05) +
  theme_map()
```
trees <- trees %>%
  st_join(hubwaystations_trees) %>%
  mutate(by_hubwaystations = !is.na(Name.y))
Now we can calculate how many trees are within 300 meters of a hubway station:

n_hubwaystations_trees <- sum(trees$by_hubwaystations)
n_hubwaystations_trees
## [1] 59110
And what percent of all trees does this represent?

n_trees <- length(trees$by_hubwaystations)
pct_hubwst_trees <-n_hubwaystations_trees / n_trees
pct_hubwst_trees



```{r, message=FALSE}
trees <- trees %>%
  st_join(trees_cc) %>%
  mutate(by_communityc = !is.na(Name.y))
```


```{r, message= FALSE}
n_trees_cc <- sum(trees$by_communityc)
n_trees_cc
```


```{r, message=FALSE}
hubways_communityc <- communityc[hubways_buffer,]
ggplot(hubways_buffer) + 
  geom_sf(data= water,fill = "lightblue", color = NA) +
   geom_sf(data= nhoods, color = "gray", fill = NA, size=0.2)+
  geom_sf() +
  geom_sf(data = hubways_communityc,
          color = "springgreen4",
          size = 0.01) +
  ggtitle("Boston, Massachusetts", subtitle = paste0("hubway stations"))+
  theme_map()
```
